Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio

### 2017-09-12 paper/structureScene.tex
rewrite abstract

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
%\title{An evaluation framework for event detection using a morphological model of acoustic scenes}
\title{Auditory Scene Similarity Retrieval and Classification with Relevance-based Quantization of Scattering Features}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
The emerging field of computational ecoacoustics aims at retrieving high-level information from the acoustic scenes recorded by some network of sensors.
In this article, we address the problem of acoustic scene classification according to the taxonomy of the DCASE 2013 challenge, both in supervised and unsupervised settings.
Traditional approaches compute a "bag-of-frames", i.e. the summary statistics of short-term audio descriptors, such as mel-frequency cepstral coefficients (MFCCs).
Although they successfully characterize static scenes with little variability in auditory content, they are unable to accurately discriminate scenes with few salient events superimposed over a textured background.
To overcome this issue, we propose a two-scale representation which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure.
We validate the use of the scattering transform for supervised acoustic scene classification by reporting a better accuracy than MFCCs when using a support vector machine (SVM) classifier.
Furthermore, we illustrate the power of the full cluster-based representation by applying it to an unsupervised acoustic scene similarity retrieval task, where we significantly outperform more traditional approaches based on summary statistics of MFCCs.
\end{abstract}

% Note that keywords are not normally used for peerreview paper.
\begin{IEEEkeywords}
unsupervised learning, classification algorithms, clustering algorithms, acoustic signal processing, wavelet transforms, audio databases, content-based retrieval, nearest neighbor searches, acoustic sensors, environmental sensors.
\end{IEEEkeywords}

@@

This works falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause, wimmer2013sampling}. Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
Consequently, few well-designed datasets are readily available for evaluation purposes.

An important problem in this field is the characterization of auditory scenes for the purposes of similarity estimation. One application of this is acoustic scene similarity retrieval, where the scenes judged ``most similar'' to a given query scene are extracted from some larger dataset. This requires the auditory scenes to be represented in some way that captures their distinctive qualities. Often, this has been achieved by bag-of-frames approaches \cite{aucouturier2007bag}, which attempts to describe the auditory scene recording using summary statistics. Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented scenes which are dynamic and contain episodic, distinct sound events that discriminate the scene from others. Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neurosciences \cite{nelken2004processing} suggest that this matches more closely the way that humans perceive acoustic scenes.

@@

Addressing this task requires first to be able to capture meaningful aspects within the signal at small time scales.
In other words, how do we represent an extract of the auditory scene signal over a sub-second interval?
A standard tool for this are the mel-frequency cepstral coefficients (MFCCs), originally developed for speech processing \cite{davis-mermelstein}, but which have found wider use in recent years in music information retrieval \cite{logan} and environmental audio processing \cite{aucouturier2007bag}.
A richer set of features is found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014}.
This is because the scattering transform cascades wavelet convolutions and pointwise nonlinearities, thus imitating the architecture of a convolutional neural network.
As a result, the scattering transform benefits from some of the assets of end-to-end learning without needing a large amount of annotated training data to converge towards a meaningful representation.

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
%\title{An evaluation framework for event detection using a morphological model of acoustic scenes}
\title{Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
The emerging field of computational ecoacoustics aims at retrieving high-level information from the acoustic scenes recorded by some network of sensors. Large amount of unlabeled data are gathered using those networks that needs to be analyzed. In order to identify which parts needs human inspection, one needs tools that can mine data and identify structural such as recurring patterns or on contrary sparse occurences. To do so, one needs to be able to measure the level of similarity among the data without strong assumptions.

State of the art approaches for computing such a similarity is the "bag-of-frames" approach, i.e. modeling the audio using the summary statistics of short-term audio descriptors, such as mel-frequency cepstral coefficients (MFCCs).
Although they successfully characterize static scenes with little variability in auditory content, they are unable to accurately discriminate scenes with few salient events superimposed over a textured background.
To overcome this issue, we propose a two-scale representation which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure that better handle the required modeling of scarce events. Experimentation performed using the Acoustic Scene Similarity framework demonstrates the superiority of the proposed approach.
\end{abstract}

% Note that keywords are not normally used for peerreview paper.
\begin{IEEEkeywords}
unsupervised learning, data mining, acoustic signal processing, wavelet transforms, audio databases, content-based retrieval, nearest neighbor searches, acoustic sensors, environmental sensors.
\end{IEEEkeywords}

@@

This works falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause, wimmer2013sampling}. %Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
%Consequently, few well-designed datasets are readily available for evaluation purposes.

An important problem in this field is the characterization of auditory scenes for the purposes of similarity estimation.

To this aim, one needs to design an algorithm to perform an acoustic scene similarity retrieval, where the scenes judged ``most similar'' to a given query scene are extracted from some larger dataset. This requires the auditory scenes to be represented in some way that captures their distinctive qualities. This has been achieved by bag-of-frames approaches \cite{aucouturier2007bag}, which attempts to describe the auditory scene recording using summary statistics. Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented scenes which are dynamic and contain episodic, distinct sound events that discriminate the scene from others. Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neurosciences \cite{nelken2004processing} suggest that this matches more closely the way that humans perceive acoustic scenes. We believe that the failure to model such distinct events is one of the reason why this approach is found to be not sufficient when considering \cite{}.

@@

Addressing this task requires first to be able to capture meaningful aspects within the signal at small time scales.
In other words, how do we represent an extract of the auditory scene signal over a sub-second interval?
A standard tool for this are the mel-frequency cepstral coefficients (MFCCs), originally developed for speech processing \cite{davis-mermelstein}, but which have found wider use in recent years in music information retrieval \cite{logan} and environmental audio processing \cite{aucouturier2007bag}.
A richer set of features is found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014}.
This is because the scattering transform cascades wavelet convolutions and pointwise nonlinearities, thus imitating the architecture of a convolutional neural network \cite{} \vl{some relevant refs on dnns 4 audio}.

For the task at at hand, the scattering features have the valuable advantage of not requiring training and thus potentially lead to wider range of application than trained. Indeed, the task at hand is the mining of yet unheard audio datasets and the properties of the audio that are of interest remains to be
